### 2019-05-22 codebook.Rmd
add codebook

---
title: "Codebook"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: 'hide'
    self_contained: true
  pdf_document:
    toc: yes
    toc_depth: 4
    latex_engine: xelatex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings during codebook generation
  message = TRUE, # show messages during codebook generation
  error = TRUE, # do not interrupt codebook generation in case of errors,
                # usually better for debugging
  echo = TRUE  # show R code
)
ggplot2::theme_set(ggplot2::theme_bw())
pander::panderOptions("table.split.table", Inf)
```

```{r prepare_codebook}
library(codebook)
library(readxl)
library(here)
library(rio)

codebook_data <- read_excel("Data_for_Analysis_of_Open_Data_and_Computational_Reproducibility_in_Registered_Reports_in_Psychology.xlsx", n_max = 62)
#Read codebook from second sheet in excel file.
dict <- read_excel("Data_for_Analysis_of_Open_Data_and_Computational_Reproducibility_in_Registered_Reports_in_Psychology.xlsx", sheet = "Codebook")

#Add variable labels to data
var_label(codebook_data) <- dict_to_list(dict)

#Add meta data
metadata(codebook_data)$name <- "Reproducibility of Registered Reports in psychology"
metadata(codebook_data)$description <- "Coded variables to examine data and code sharing and the computational reproducibility for Registered Reports in psychology"
metadata(codebook_data)$creator <- "Pepijn Obels, Daniel Lakens, Nicholas A. Coles, Jaroslav Gottfried"
metadata(codebook_data)$temporalCoverage <- "Spring 2018"

#Export codebook
rio::export(codebook_data, "reproducing_registered_reports_codebook.rds") # to R data structure file
```

```{r codebook}
codebook(codebook_data)
```